\begin{document}

\title{The Odes of Solomon}

\chapter{1}

\par 1 The Lord is on my head like a crown, and I shall not be without Him.
\par 2 They wove for me a crown of truth, and it caused thy branches to bud in me.
\par 3 For it is not like a withered crown which buddeth not: but thou livest upon my head, and thou hast blossomed upon my head.
\par 4 Thy fruits are full-grown and perfect, they are full of thy salvation.



\chapter{2}

\par \textit{(No part of this Ode has ever been identified.)}

\chapter{3}

\par \textit{The first words of this Ode have disappeared.}

\par 1 .    .    .   I put on:
\par 2 And his members are with him. And on them do I stand, and He loves me:
\par 3 For I should not have known how to love the Lord, if He had not loved me.
\par 4 For who is able to distinguish love, except the one that is loved?
\par 5 I love the Beloved, and my soul loves Him:
\par 6 And where His rest is, there also am I;
\par 7 And I shall be no stranger, for with the Lord Most High and Merciful there is no grudging.
\par 8 I have been united to I-run, for the Lover has found the Beloved,
\par 9 And because I shall love Him that, is the Son, I shall become a son;
\par 10 For he that is joined to Him that is immortal, will also himself become immortal;
\par 11 And he who has pleasure in the Living One, will become living.
\par 12 This is the Spirit of the Lord, which doth not lie, which teacheth the sons of men to know His ways.
\par 13 Be wise and understanding and vigilant. Hallelujah.

\chapter{4}

\par \textit{This Ode is important because of the historical allusion with which it commences. This may refer to the closing of the temple at Leontopolis in Egypt which would date this writing about 73 A. D.}

\par 1 No man, O my God, changeth thy holy place;
\par 2 And it is not (possible) that he should change it and put it in another place: because he hath no power over it:
\par 3 For thy sanctuary thou hast designed before thou didst make (other) places:
\par 4 That which is the older shall not be altered by those that are younger than itself.
\par 5 Thou has given thy heart, O Lord, to thy believers: never wilt thou fail, nor be without fruits:
\par 6 For one hour of thy Faith is more precious than all days and years.
\par 7 For who is there that shall put on thy grace, and be hurt?
\par 8 For thy seal is known: and thy creatures know it: and thy (heavenly) hosts possess it: and the elect archangels are clad with it.
\par 9 Thou hast given us thy fellowship: it was not that thou wast in need of us: but that we are in need of thee:
\par 10 Distill thy dews upon us and open thy rich fountains that pour forth to us milk and honey:
\par 11 For there is no repentance with thee that thou shouldest repent of anything that thou hast promised:
\par 12 And the end was revealed before thee: for what thou gavest, thou gavest freely:
\par 13 So that thou mayest, not draw them back and take them again:
\par 14 For all was revealed before thee as God, and ordered from the beginning before thee: and thou, O God, hast made all things. Hallelujah.

\chapter{5}

\par \textit{This Ode has strangely appeared in a speech by Salome in another ancient work called the Pistis Sophia.}

\par 1 I will give thanks unto thee, O Lord, because I love thee;
\par 2 O Most High, thou wilt not forsake me, for thou art my hope:
\par 3 Freely I have received thy grace, I shall live thereby:
\par 4 My persecutors will come and not see me:
\par 5 A cloud of darkness shall fall on their eyes; and an air of thick gloom shall darken them:
\par 6 And they shall have no light to see: they may not take hold upon me.
\par 7 Let their counsel become thick darkness, and what I have cunningly devised, let it return upon their own heads:
\par 8 For they have devised a counsel, and it did not succeed:
\par 9 For my hope is upon the Lord, and I will not fear, and because the Lord is my salvation, I will not fear:
\par 10 And He is as a garland on my head and I shall not be moved; even if everything should be shaken, I stand firm;
\par 11 And if all things visible should perish, I shall not die; because the Lord is with me and I am with Him. Hallelujah.

\chapter{6}

\par \textit{First century universalism is revealed in an interesting way in verse 10.}

\par 1 As the hand moves over the harp, and the strings speak.
\par 2 So speaks in my members the Spirit of the Lord, and I speak by His love.
\par 3 For it destroys what is foreign, and everything that is bitter:
\par 4 For thus it was from the beginning and will be to the end, that nothing should be His adversary, and nothing should stand up against Him.
\par 5 The Lord has multiplied the knowledge of Himself, and is zealous that these things should be known, which by His grace have been given to us.
\par 6 And the praise of His name He gave us: our spirits praise His holy Spirit.
\par 7 For there went forth a stream and became a river great and broad;
\par 8 For it flooded and broke up everything and it brought (water) to the Temple:
\par 9 And the restrainers of the children of men were not able to restrain it, nor the arts of those whose business it is to restrain waters;
\par 10 For it spread over the face of the whole earth, and filled everything: and all the thirsty upon earth were given to drink of it;
\par 11 And thirst was relieved and quenched: for from the Most High the draught was given.
\par 12 Blessed then are the ministers of that draught who are entrusted with that water of His:
\par 13 They have assuaged the dry lips, and the will that had fainted they have raised up;
\par 14 And souls that were near departing they have caught back from death:
\par 15 And limbs that had fallen they straightened and set up:
\par 16 They gave strength for their feebleness and light to their eyes:
\par 17 For everyone knew them in the Lord, and they lived by the water of life for ever. Hallelujah.



\chapter{7}

\par \textit{A wonderfully, simple and joyful psalm on the Incarnation.}

\par 1 As the impulse of anger against evil, so is the impulse of joy over what is lovely, and brings in of its fruits without restraint:
\par 2 My joy is the Lord and my impulse is toward Him: this path of mine is excellent:
\par 3 For I have a helper, the Lord.
\par 4 He hath caused me to know Himself, without grudging, by His simplicity: His kindness has humbled His greatness.
\par 5 He became like me, in order that I might receive Him:
\par 6 He was reckoned like myself in order that I might put Him on;
\par 7 And I trembled not when I saw Him: because He was gracious to me:
\par 8 Like my nature He became that I might learn Him and like my form, that I might not turn back from Him:
\par 9 The Father of knowledge is the word of knowledge:
\par 10 He who created wisdom is wiser than His works:
\par 11 And He who created me when yet I was not knew what I should do when I came into being:
\par 12 Wherefore He pitied me in His abundant grace: and granted me to ask from Him and to receive from His sacrifice:
\par 13 Because He it is that is incorrupt, the fulness of the ages and the Father of them.
\par 14 He hath given Him to be seen of them that are His, in order that they may recognize Him that made them: and that they might not suppose that they came of themselves:
\par 15 For knowledge He hath appointed as its way, He hath widened it and extended it; and brought to all perfection;
\par 16 And set over it the traces of His light, and I walked therein from the beginning even to the end.
\par 17 For by Him it was wrought, and He was resting in the Son, and for its salvation He will take hold of everything;
\par 18 And the Most High shall be known in His Saints, to announce to those that have songs of the coming of the Lord;
\par 19 That they may go forth to meet Him, and may sing to Him with joy and with the harp of many tones:
\par 20 The seers shall come before Him and they shall be seen before Him,
\par 21 And they shall praise the Lord for His love: because He is near and beholdeth.
\par 22 And hatred shall be taken from the earth, and along with jealousy it shall be drowned:
\par 23 For ignorance hath been destroyed, because the knowledge of the Lord hath arrived.
\par 24 They who make songs shall sing the grace of the Lord Most High;
\par 25 And they shall bring their songs, and their heart shall be like the day: and like the excellent beauty of the Lord their pleasant song:
\par 26 And there shall neither be anything that breathes without knowledge, nor any that is dumb:
\par 27 For He hath given a mouth to His creation, to open the voice of the mouth towards Him, to praise Him:
\par 28 Confess ye His power, and show forth His grace. Hallelujah.

\chapter{8}

\par \textit{Note the sudden transition from the person of the Psalmist to the person of the Lord (v. 10). This is like the canonical Psalter in style.}

\par 1 Open ye, open ye your hearts to the exultation of the Lord:
\par 2 And let your love be multiplied from the heart and even to the lips,
\par 3 To bring forth fruit to the Lord [fruit], holy [fruit], and to talk with watchfulness in His light.
\par 4 Rise up, and stand erect, ye who sometime were brought low:
\par 5 Tell forth ye who were in silence, that your mouth hath been opened.
\par 6 Ye, therefore, that were despised, be henceforth lifted up, because your righteousness hath been exalted.
\par 7 For the right hand of the Lord is with you: and He is your helper:
\par 8 And peace was prepared for you, before ever your war was.
\par 9 Hear the word of truth, and receive the knowledge of the Most High.
\par 10 Your flesh has not known what I am saying to you neither have your hearts known what I am showing to you.
\par 11 Keep my secret, ye who are kept by it:
\par 12 Keep my faith, ye who are kept by it.
\par 13 And understand my knowledge, ye who know me in truth.
\par 14 Love me with affection, ye who love:
\par 15 For I do not turn away my face from them that are mine;
\par 16 For I know them, and before they came into being I took knowledge of them, and on their faces I set my seal:
\par 17 I fashioned their members: my own breasts I prepared for them, that they might drink my holy milk and live thereby.
\par 181 took pleasure in them and am not ashamed of them:
\par 19 For my workmanship are they and the strength of my thoughts:
\par 20 Who then shall rise up against my handiwork, or who is there that is not subject to them?
\par 21 I willed and fashioned mind and heart: and they are mine, and by my own right hand I set my elect ones:
\par 22 And my righteousness goeth before them and they shall not be deprived of my name, for it is with them.
\par 23 Ask, and abound and abide in the love of the Lord,
\par 24 And yet beloved ones in the Beloved: those who are kept, in Him that liveth:
\par 25 And they that are saved in Him that was saved;
\par 26 And ye shall be found incorrupt in all ages to the name of your Father. Hallelujah.

\chapter{9}

\par \textit{We shall never know surely whether the wars referred to here are spiritual or actual outward wars.}

\par 1 Open your ears and I will speak to you. Give me your souls that I may also give you my soul,
\par 2 The word of the Lord and His good pleasures, the holy thought which He has devised concerning his Messiah.
\par 3 For in the will of the Lord is your salvation, and His thought is everlasting life; and your end is immortality.
\par 4 Be enriched in God the Father, and receive the thought of the Most High.
\par 5 Be strong and be redeemed by His grace.
\par 6 For I announce to you peace, to you His saints;
\par 7 That none of those who hear may fall in war, and that those again who have known Him may not perish, and that those who receive may not be ashamed.
\par 8 An everlasting crown for ever is Truth. Blessed are they who set it on their heads:
\par 9 A stone of great price is it; and there have been wars on account of the crown.
\par 10 And righteousness hath taken it and hath given it to you.
\par 11 Put on the crown in the true covenant of the Lord.
\par 12 And all those who have conquered shall be written in His book.
\par 13 For their book is victory which is yours. And she (Victory) sees you before her and wills that you shall be saved. Hallelujah.

\chapter{10}

\par \textit{A vigorous little Ode in which Christ Himself is the speaker.}

\par 1 The Lord hath directed my mouth by His word: and He hath opened my heart by His light: and He hath caused to dwell in me His deathless life;
\par 2 And gave me that I might speak the fruit of His peace:
\par 3 To convert the souls of them who are willing to come to Him; and to lead captive a good captivity for freedom.
\par 4 I was strengthened and made mighty and took the world captive;
\par 5 And it became to me for the praise of the Most High, and of God my Father.
\par 6 And the Gentiles were gathered together who were scattered abroad.
\par 7 And I was unpolluted by my love for them, because they confessed me in high places: and the traces of the light were set upon their heart:
\par 8 And they walked in my life and were saved and became my people for ever and ever. Hallelujah.

\chapter{11}

\par \textit{A beautiful sketch of Paradise regained and the blessedness of those who have returned to the privileges of the fallen Adam.}

\par 1 My heart was cloven and its flower appeared; and grace sprang up in it: and it brought forth fruit to the Lord,
\par 2 For the Most High clave my heart by His Holy Spirit and searched my affection towards Him: and filled me with His love.
\par 3 And His opening of me became my salvation; and I ran in His way in His peace, even in the way of truth:
\par 4 From the beginning and even to the end I acquired His knowledge:
\par 5 And I was established upon the rock of truth, where He had set me up:
\par 6 And speaking waters touched my lips from the fountain of the Lord plenteously:
\par 7 And I drank and was inebriated with the living water that doth not die;
\par 8 And my inebriation was not one without knowledge, but I forsook vanity and turned to the Most High my God,
\par 9 And I was enriched by His bounty, and I forsook the folly which is diffused over the earth; and I stripped it off and cast it from me:
\par 10 And the Lord renewed me in His raiment, and possessed me by His light, and from above He gave me rest in incorruption;
\par 11 And I became like the land which blossoms and rejoices in its fruits:
\par 12 And the Lord was like the sun shining on the face of the land;
\par 13 He lightened my eyes, and my face received the dew; and my nostrils enjoyed the pleasant odour of the Lord;
\par 14 And He carried me to His Paradise; where is the abundance of the pleasure of the Lord;
\par 15 And I worshipped the Lord on account of His glory; and I said, Blessed, O Lord, are they who are planted in thy land! and those who have a place in thy Paradise;
\par 16 And they grow by the fruits of the trees. And they have changed from darkness to light.
\par 17 Behold! all thy servants are fair, who do good works, and turn away from wickedness to the pleasantness that is thine:
\par 18 And they have turned back the bitterness of the trees from them, when they were planted in thy, land;
\par 19 And everything became like a relic of thyself, and memorial for ever of thy faithful works.
\par 20 For there is abundant room in thy Paradise, and nothing is useless therein;
\par 21 But everything is filled with fruit; glory be to thee, O God, the delight of Paradise for ever. Hallelujah.

\chapter{12}

\par \textit{An exceptionally high level of spiritual thought.}

\par 1 He hath filled me with words of truth; that I may speak the same;
\par 2 And like the flow of waters flows truth from my mouth, and my lips show forth His fruit.
\par 3 And He has caused His knowledge to abound in me, because the mouth of the Lord is the true Word, and the door of His light;
\par 4 And the Most High hath given it to His words, which are the interpreters of His own beauty, and the repeaters of His praise, and the confessors of His counsel, and the heralds of His thought, and the chasteners of His servants.
\par 5 For the swiftness of the Word is inexpressible, and like its expression is its swiftness and force;
\par 6 And its course knows no limit. Never doth it fail, but it stands sure, and it knows not descent nor the way of it.
\par 7 For as its work is, so is its end: for it is light and the dawning of thought;
\par 8 And by it the worlds talk one to the other; and in the Word there were those that were silent;
\par 9 And from it came love and concord; and they spake one to the other whatever was theirs; and they were penetrated by the Word;
\par 10 And they knew Him who made them, because they were in concord; for the mouth of the Most High spake to them; and His explanation ran by means of it:
\par 11 For the dwelling-place of the Word is man: and its truth is love.
\par 12 Blessed are they who by means thereof have understood everything, and have known the Lord in His truth. Hallelujah.

\chapter{13}

\par \textit{A strange little Ode.}

\par 1 Behold! the Lord is our mirror: open the eyes and see them in Him: and learn the manner of your face:
\par 2 And tell forth praise to His spirit: and wipe off the filth from your face: and love His holiness and clothe yourselves therewith:
\par 3 And be without stain at all times before Him. Hallelujah.

\chapter{14}

\par \textit{This Ode is as beautiful in style as the canonical Psalter.}

\par 1 As the eyes of a son to his father, so are my eyes, O Lord, at all times towards thee.
\par 2 For with thee are my consolations and my delight.
\par 3 Turn not away thy mercies from me, O Lord: and take not thy kindness from me.
\par 4 Stretch out to me, O Lord, at all times thy right hand: and be my guide even unto the end, according to thy good pleasure.
\par 5 Let me be well-pleasing before thee, because of thy glory and because of thy name:
\par 6 Let me be preserved from evil, and let thy meekness, O Lord, abide with me, and the fruits of thy love.
\par 7 Teach me the Psalms of thy truth, that I may bring forth fruit in thee:
\par 8 And open to me the harp of thy Holy Spirit, that with all its notes I may praise thee, O Lord.
\par 9 And according to the multitude of thy tender mercies, so thou shalt give to me; and hasten to grant our petitions; and thou art able for all our needs. Hallelujah.



\chapter{15}

\par \textit{One of the loveliest Odes in this unusual collection.}

\par 1 As the sun is the joy to them that seek for its daybreak, so is my joy the Lord;
\par 2 Because He is my Sun and His rays have lifted me up; and His light hath dispelled all darkness from my face.
\par 3 In Him I have acquired eyes and have seen His holy day:
\par 4 Ears have become mine and I have heard His truth.
\par 5 The thought of knowledge hath been mine, and I have been delighted through Him.
\par 6 The way of error I have left, and have walked towards Him and have received salvation from Him, without grudging.
\par 7 And according to His bounty He hath given to me, and according to His excellent beauty He hath made me.
\par 8 I have put on incorruption through His name: and have put oft corruption by His grace.
\par 9 Death hath been destroyed before my face: and Sheol hath been abolished by my word:
\par 10 And there hath gone up deathless life in the Lord's land,
\par 11 And it hath been made known to His faithful ones, and hath been given without stint to all those that trust in Him. Hallelujah.

\chapter{16}

\par \textit{The beauty of God's creation.}

\par 1 As the work of the husbandman is the ploughshare: and the work of the steersman is the guidance of the ship:
\par 2 So also my work is the Psalm of the Lord: my craft and my occupation are in His praises:
\par 3 Because His love bath nourished my heart, and even to my lips His fruits He poured out.
\par 4 For my love is the Lord, and therefore I will sing unto Him:
\par 5 For I am made strong in His praise, and I have faith in Him.
\par 6 I will open my mouth and His spirit will utter in me the glory of the Lord and His beauty; the work of His hands and the operation of His fingers:
\par 7 The multitude of His mercies and the strength of His word.
\par 8 For the word of the Lord searches out all things, both the invisible and that which reveals His thought;
\par 9 For the eye sees His works, and the ear hears His thought;
\par 10 He spread out the earth and He settled the waters in the sea:
\par 11 He measured the heavens and fixed the stars: and He established the creation and set it up:
\par 12 And He rested from His works:
\par 13 And created things run in their courses, and do their works:
\par 14 And they know not how to stand and be idle; and His heavenly hosts are subject to His word.
\par 15 The treasure-chamber of the light is the sun, and the treasury of the darkness is the night:
\par 16 And He made the sun for the day that it may be bright, but night brings darkness over the face of the land;
\par 17 And their alternations one to the other speak the beauty of God:
\par 18 IS And there is nothing that is without the Lord; for He was before any thing came into being:
\par 19 And the worlds were made by His word, and by the thought of His heart. Glory and honour to His name. Hallelujah.

\chapter{17}

\par \textit{A peculiar change of personality, scarcely realized until the return from it in the last verse.}

\par 1 I was crowned by my God: my crown is living:
\par 2 And I was justified in my Lord: my incorruptible salvation is He.
\par 3 I was loosed from vanity, and I was not condemned:
\par 4 The choking bonds were cut off by her hands: I received the face and the fashion of a new person: and I walked in it and was saved;
\par 5 And the thought of truth led me on. And I walked after it and did not wander:
\par 6 And all that have seen me were amazed: and I was regarded by them as a strange person:
\par 7 And He who knew and brought me up is the Most High in all His perfection. And He glorified me by His kindness, and raised my thoughts to the height of His truth.
\par 8 And from thence He gave me the way of His precepts and I opened the doors that were closed,
\par 9 And brake in pieces the bars of iron; but my iron melted and dissolved before me;
\par 10 Nothing appeared closed to me: because I was the door of everything.
\par 11 And I went over all my bondmen to loose them; that I might not leave any man bound or binding:
\par 12 And I imparted my knowledge without grudging: and my prayer was in my love:
\par 13 And I sowed my fruits in hearts, and transformed them into myself: and they received my blessing and lived;
\par 14 And they were gathered to me and were saved; because they were to me as my own members and I was their head. Glory to thee our head, the Lord Messiah. Hallelujah.

\chapter{18}

\par \textit{A man who had a spiritual experience brings a message.}

\par 1 My heart was lifted up in the love of the Most High and was enlarged: that I might praise Him for His name's sake.
\par 2 My members were strengthened that they might not fall from His strength.
\par 3 Sicknesses removed from my body, and it stood to the Lord by His will. For His kingdom is true.
\par 4 O Lord, for the sake of them that are deficient do not remove thy word from me!
\par 5 Neither for the sake of their works do thou restrain from me thy perfection!
\par 6 Let not the luminary be conquered by the darkness; nor let truth flee away from falsehood.
\par 7 Thou wilt appoint me to victory; our Salvation is thy right hand. And thou wilt receive men from all quarters.
\par 8 And thou wilt preserve whosoever is held in evils:
\par 9 Thou art my God. Falsehood and death are not in thy mouth:
\par 10 For thy will is perfection; and vanity thou knowest not,
\par 11 Nor does it know thee.
\par 12 And error thou knowest not,
\par 13 Neither does it know thee.
\par 14 And ignorance appeared like a blind man; and like the foam of the sea,
\par 15 And they supposed of that vain thing that it was something great;
\par 16 And they too came in likeness of it and became vain; and those have understood who have known and meditated;
\par 17 And they have not been corrupt in their imagination; for such were in the mind of the Lord;
\par 18 And they mocked at them that were walking in error;
\par 19 And they spake truth from the inspiration which the Most High breathed into them; Praise and great comeliness to His name Hallelujah.

\chapter{19}

\par \textit{Fantastic and not in harmony with the other Odes. The reference to a painless Virgin Birth is notable.}

\par 1 A cup of milk was offered to me: and I drank it in the sweetness of the delight of the Lord.
\par 2 The Son is the cup, and He who was milked is the Father:
\par 3 And the Holy Spirit milked Him: because His breasts were full, and it was necessary for Him that His milk should be sufficiently released;
\par 4 And the Holy Spirit opened His bosom and mingled the milk from the two breasts of the Father; and gave the mixture to the world without their knowing:
\par 5 And they who receive in its fulness are the ones on the right hand.
\par 6 The Spirit opened the womb of the Virgin and she received conception and brought forth; and the Virgin became a Mother with many mercies;
\par 7 And she travailed and brought forth a Son, without incurring pain;
\par 8 And because she was not sufficiently prepared, and she had not sought a midwife (for He brought her to bear) she brought forth, as if she were a man, of her own will;
\par 9 And she brought Him forth openly, and acquired Him with great dignity,
\par 10 And loved Him in His swaddling clothes and guarded Him kindly, and showed Him in Majesty. Hallelujah.

\chapter{20}

\par \textit{A mixture of ethics and mysticism; of the golden rule and the tree of life.}

\par 1 I am a priest of the Lord, and to Him I do priestly service: and to Him I offer the sacrifice of His thought.
\par 2 For His thought is not like the thought of the world nor the thought of the flesh, nor like them that serve carnally.
\par 3 The sacrifice of the Lord is righteousness, and purity of heart and lips.
\par 4 Present your reins before Him blamelessly: and let not thy heart do violence to heart, nor thy soul to soul.
\par 5 Thou shalt not acquire a stranger by the price of thy silver, neither shalt thou seek to devour thy neighbour,
\par 6 Neither shalt thou deprive him of the covering of his nakedness.
\par 7 But put on the grace of the Lord without stint; and come into His Paradise and make thee a garland from its tree,
\par 8 And put it on thy head and be glad; and recline on His rest, and glory shall go before thee,
\par 9 And thou shalt receive of His kindness and of His grace; and thou shalt be flourishing in truth in the praise of His holiness. Praise and honour be to His name. Hallelujah.

\chapter{21}

\par \textit{A remarkable explanation of the “coats of skin” in the third chapter of Genesis.}

\par 1 My arms I lifted up to the Most High, even to the grace of the Lord: because He had cast off my bonds from me: and my Helper had lifted me up to His grace and to His salvation:
\par 2 And I put off darkness and clothed myself with light,
\par 3 And my soul acquired a body free from sorrow or affliction or pains.
\par 4 And increasingly helpful to me was the thought of the Lord, and His fellowship in incorruption:
\par 5 And I was lifted up in His light; and I served before Him,
\par 6 And I became near to Him, praising and confessing Him;
\par 7 My heart ran over and was found in my mouth: and it arose upon my lips; and the exultation of the Lord increased on my face, and His praise likewise. Hallelujah.

\chapter{22}

\par \textit{Like the Psalms of David in their exultation because of freedom.}

\par 1 He who brought me down from on high, also brought me up from the regions below;
\par 2 And He who gathers together the things that are betwixt is He also who cast me down:
\par 3 He who scattered my enemies had existed from ancient and my adversaries:
\par 4 He who gave me authority over bonds that I might loose them;
\par 5 He that overthrew by my hands the dragon with seven heads: and thou hast set me over his roots that I might destroy his seed.
\par 6 Thou wast there and didst help me, and in every place thy name was a rampart me
\par 7 Thy right hand destroyed his wicked poison; and thy hand levelled the way for those who believe in thee.
\par 8 And thou didst choose them from the graves and didst separate them from the dead.
\par 9 Thou didst take dead bones and didst cover them with bodies.
\par 10 They were motionless, and thou didst give them energy for life.
\par 11 Thy way was without corruption, and thy face; thou didst bring thy world to corruption: that everything might be dissolved, and then renewed,
\par 12 And that the foundation for everything might be thy rock: and on it thou didst build thy kingdom; and it became the dwelling place of the saints. Hallelujah.

\chapter{23}

\par \textit{The reference to the sealed document sent by God is one of the great mysteries of the collection.}

\par 1 Joy is of the saints! and who shall put it on, but they alone?
\par 2 Grace is of the elect! and who shall receive it except those who trust in it from the beginning?
\par 3 Love is of the elect? And who shall put it on except those who have possessed it from the beginning?
\par 4 Walk ye in the knowledge of the Most High without grudging: to His exultation and to the perfection of His knowledge.
\par 5 And His thought was like a letter; His will descended from on high, and it was sent like an arrow which is violently shot from the bow:
\par 6 And many hands rushed to the letter to seize it and to take and read it:
\par 7 And it escaped their fingers and they were affrighted at it and at the seal that was upon it.
\par 8 Because it was not permitted to them to loose its seal: for the power that was over the seal was greater than they.
\par 9 But those who saw it went after the letter that they might know where it would alight, and who should read it and who should hear it.
\par 10 But a wheel received it and came over it:
\par 11 And there was with it a sign of the Kingdom and of the Government:
\par 12 And everything which tried to move the wheel it mowed and cut down:
\par 13 And it gathered the multitude of adversaries, and bridged the rivers and crossed over and rooted up many forests and made a broad path.
\par 14 The head went down to the feet, for down to the feet ran the wheel, and that which was a sign upon it.
\par 15 The letter was one of command, for there were included in it all districts;
\par 16 And there was seen at its head, the head which was revealed even the Son of Truth from the Most High Father,
\par 17 And He inherited and took possession of everything. And the thought of many was brought to nought.
\par 18 And all the apostates hasted and fled away. And those who persecuted and were enraged became extinct.
\par 19 And the letter was a great volume, which was wholly written by the. finger of God:
\par 20 And the name of the Father was on it, and of the Son and of the Holy Spirit, to rule for ever and ever. Hallelujah.

\chapter{24}

\par \textit{The mention of the Dove refers to a lost Gospel to which there are rare references in ancient writings.}

\par 1 The Dove fluttered over the Messiah, because He was her head; and she sang over Him and her voice was heard:
\par 2 And the inhabitants were afraid and the sojourners were moved:
\par 3 The birds dropped their wings and all creeping things died in their holes: and the abysses were opened which had been hidden; and they cried to the Lord like women in travail:
\par 4 And no food was given to them, because it did not belong to them;
\par 5 And they sealed up the abysses with the seal of the Lord. And they perished, in the thought, those that had existed from ancient times;
\par 6 For they were corrupt from the beginning; and the end of their corruption was life:
\par 7 And every one of them that was imperfect perished: for it was not possible to give them a word that they might remain:
\par 8 And the Lord destroyed the imaginations of all them that had not the truth with them.
\par 9 For they who in their hearts were lifted up were deficient in wisdom, and so they were rejected, because the truth was not with them.
\par 10 For the Lord disclosed His way, and spread abroad His grace: and those who understood it, know His holiness. Hallelujah.

\chapter{25}

\par \textit{Back again to personal experience.}

\par 1 I was rescued from my bonds and unto thee, my God, I fled:
\par 2 For thou art the right hand of my Salvation and my helper.
\par 3 Thou hast restrained those that rise up against me,
\par 4 And I shall see him no more: because thy face was with me, which saved me by thy grace.
\par 5 But I was despised and rejected in the eye of many: and I was in their eyes like lead,
\par 6 And strength was mine from thyself and help.
\par 7 Thou didst set me a lamp at my right hand and at my left: and in me there shall be nothing that is not bright:
\par 8 And I was clothed with the covering of thy Spirit, and thou didst remove from me my raiment of skin;
\par 9 For thy right hand lifted me up and removed sickness from me:
\par 10 And I became mighty in: the truth, and holy by thy righteousness; and all my adversaries were afraid of me;
\par 11 And I became admirable by the name of the Lord, and was justified by His gentleness, and His rest is for ever and ever. Hallelujah.

\chapter{26}

\par \textit{Remarkable praise.}

\par 1 I poured out praise to the Lord, for I am His:
\par 2 And I will speak His holy song, for my heart is with Him.
\par 3 For His harp is in my hands, and the Odes of His rest shall not be silent.
\par 4 I will cry unto him from my whole heart: I will praise and exalt Him with all my members.
\par 5 For from the east and even to the west is His praise:
\par 6 And from the south and even to the north is the confession of Him:
\par 7 And from the top of the hills to their utmost bound is His perfection.
\par 8 Who can write the Psalms of the Lord, or who read them?
\par 9 Or who can train his soul for life, that his soul may be saved,
\par 10 Or who can rest on the Most High, so that with His mouth he may speak?
\par 11 Who is able to interpret the wonders of the Lord?
\par 12 For he who could interpret would be dissolved and would become that which is interpreted.
\par 13 For it suffices to know and to rest: for in rest the singers stand,
\par 14 Like a river which has an abundant fountain, and flows to the help of them that seek it. Hallelujah.

\chapter{27}

\par \textit{The human body makes a cross when a man stands erect in prayer with arms outstretched.}

\par 1 I stretched out my hands and sanctified my Lord:
\par 2 For the extension of my hands is His sign:
\par 3 And my expansion is the upright tree [or cross].



\chapter{28}

\par \textit{This Ode is a musical gem.}

\par 1 As the wings of doves over their nestlings; and the mouth of their nestlings towards their mouths.
\par 2 So also are the wings of the Spirit over my heart:
\par 3 My heart is delighted and exults: like the babe who exults in the womb of his mother:
\par 4 I believed; therefore I was at rest; for faithful is He in whom I have believed:
\par 5 He has richly blessed me and my head is with Him: and the sword shall not divide me from Him, nor the scimitar;
\par 6 For I am ready before destruction comes; and I have been set on His immortal pinions:
\par 7 And He showed me His sign: forth and given me to drink, and from that life is the spirit within me, and it cannot die, for it lives.
\par 8 They who saw me marvelled at me, because I was persecuted, and they supposed that I was swallowed up: for I seemed to them as one of the lost;
\par 9 And my oppression became my salvation; and I was their reprobation because there was no zeal in me;
\par 10 Because I did good to every man I was hated,
\par 11 And they came round me like mad dogs, who ignorantly attack their masters,
\par 12 For their thought is corrupt and their understanding perverted.
\par 13 But I was carrying water in my right hand, and their bitterness I endured by my sweetness;
\par 14 And I did not perish, for I was not their brother nor was my birth like theirs.
\par 15 And they sought for my death and did not find it: for I was older than the memorial of them;
\par 16 And vainly did they make attack upon me and those who, without reward, came after me:
\par 17 They sought to destroy the memorial of him who was before them.
\par 18 For the thought of the Most High cannot be anticipated; and His heart is superior to all wisdom. Hallelujah.

\chapter{29}

\par \textit{Again reminiscent of the Psalms, of David.}

\par 1 The Lord is my hope: in Him I shall not be confounded.
\par 2 For according to His praise He made me, and according to His goodness even so He gave unto me:
\par 3 And according to His mercies He exalted me: and according to His excellent beauty He set me on high:
\par 4 And brought me up out of the depths of Sheol: and from the mouth of death He drew me:
\par 5 And thou didst lay my enemies low, and He justified me by His grace.
\par 6 For I believed in the Lord's Messiah: and it appeared to me that He is the Lord;
\par 7 And He showed him His sign: and He led me by His light, and gave me the rod of His power;
\par 8 That I might subdue the imaginations of the peoples; and the power of the men of might to bring them low:
\par 9 To make war by His word, and to take victory by His power.
\par 10 And the Lord overthrew my enemy by His word: and he became like the stubble which the wind carries away;
\par 11 And I gave praise to the Most High because He exalted me His servant and the son of His handmaid. Hallelujah.

\chapter{30}

\par \textit{An invitation to the thirsty.}

\par 1 Fill ye waters for yourselves from the living fountain of the Lord, for it is opened to you:
\par 2 And come all ye thirsty, and take the draught; and rest by the fountain of the Lord.
\par 3 For fair it is and pure and gives rest to the soul. Much more pleasant are its waters than honey;
\par 4 And the honeycomb of bees is not to be compared with it.
\par 5 For it flows forth from the lips of the Lord and from the heart of the Lord is its name.
\par 6 And it came infinitely and invisibly: and until it was set in the midst they did not know it:
\par 7 Blessed are they who have drunk therefrom and have found rest thereby. Hallelujah.

\chapter{31}

\par \textit{A song that Marcus Aurelius might have known when he said “Be like the promontory against which the waves continually break.”}

\par 1 The abysses were dissolved before the Lord: and darkness was destroyed by His appearance:
\par 2 Error went astray and perished at His hand: and folly found no path to walk in, and was submerged by the truth of the Lord.
\par 3 He opened His mouth and spake grace and joy: and He spake a new song of praise to His name:
\par 4 And He lifted up His voice to the Most High, and offered to Him the sons that were with Him.
\par 5 And His face was justified, for thus His holy Father had given to Him.
\par 6 Come forth, ye that have been afflicted and receive joy, and possess your souls by His grace; and take to you immortal life.
\par 7 And they made me a debtor when I rose up, me who had been a debtor: and they divided my spoil, though nothing was due to them.
\par 8 But I endured and held my peace and was silent, as if not, moved by them.
\par 9 But I stood unshaken like a firm rock which is beaten by the waves and endures.
\par 10 And I bore their bitterness for humility's sake:
\par 11 In order, that I might redeem my people, and inherit it and that I might not make void my promises to the fathers, to whom I promised the salvation of their seed. Hallelujah.

\chapter{32}

\par \textit{Joy and light.}

\par 1 To the blessed there is joy from their hearts, and light from Him that dwells in them:
\par 2 And words from the Truth, who was self-originate: for He is strengthened by the holy power of the Most High: and He is unperturbed for ever and ever. Hallelujah.

\chapter{33}

\par \textit{A virgin stands and proclaims (v. 5).}

\par 1 Again Grace ran and forsook corruption, and came down in Him to bring it to nought;
\par 2 And He destroyed perdition from before Him, and devastated all its order;
\par 3 And He stood on a lofty summit and uttered His voice from one end of the earth to the other:
\par 4 And drew to Him all those who obeyed Him; and there did not appear as it were an evil person.
\par 5 But there stood a perfect virgin who was proclaiming and calling and saying,
\par 6 O ye sons of men, return ye, and ye daughters of men, come ye:
\par 7 And forsake the ways of that corruption and draw near unto me, and I will enter into you, and will bring you forth from perdition,
\par 8 And make you wise in the ways of truth: that you be not destroyed nor perish:
\par 9 Hear ye me and be redeemed. For the grace of God I am telling among you: and by my means you shall be redeemed and become blessed.
\par 10 I am your judge; and they who have put me on shall not be injured: but they shall possess the new world that is incorrupt:
\par 11 My chosen ones walk in me, and my ways I will make known to them that seek me, and I will make them trust in my name. Hallelujah.



\chapter{34}

\par \textit{True poetry—pure and simple.}

\par 1 No way is hard where there is a simple heart.
\par 2 Nor is there any wound where the thoughts are upright:
\par 3 Nor is there any storm in the depth of the illuminated thought:
\par 4 Where one is surrounded on every side by beauty, there is nothing that is divided.
\par 5 The likeness of what is below is that which is above; for everything is above: what is below is nothing but the imagination of those that are without knowledge.
\par 6 Grace has been revealed for your salvation. Believe and live and be saved. Hallelujah.

\chapter{35}

\par \textit{“No cradled child more softly lies than I: come soon, eternity.”}

\par 1 The dew of the Lord in quietness He distilled upon me:
\par 2 And the cloud of peace He caused to rise over my head, which guarded me continually;
\par 3 It was to me for salvation: everything was shaken and they were affrighted;
\par 4 And there came forth from them a smoke and a judgment; and I was keeping quiet in the order of the Lord:
\par 5 More than shelter was He to me, and more than foundation.
\par 6 And I was carried like a child by his mother: and He gave me milk, the dew of the Lord:
\par 7 And I grew great by His bounty, and rested in His perfection,
\par 8 And I spread out my hands in the lifting up of my soul: and I was made right with the Most High, and I was redeemed with Him. Hallelujah.

\chapter{36}

\par \textit{Theologians have never agreed on an explanation of this perplexing Ode.}

\par 1 I rested in the Spirit of the Lord: and the Spirit raised me on high:
\par 2 And made me stand on my feet in the height of the Lord, before His perfection and His glory, while I was praising Him by the composition of His songs.
\par 3 The Spirit brought me forth before the face of the Lord: and, although a son of man, I was named the Illuminate, the Son of God:
\par 4 While I praised amongst the praising ones, and great was I amongst the mighty ones.
\par 5 For according to the greatness of the Most High, so He made me: and like His own newness He renewed me; and He anointed me from His own perfection:
\par 6 And I became one of His Neighbours; and my mouth was opened, like a cloud of dew,
\par 7 And my heart poured out as it were a gushing stream of righteousness,
\par 8 And my access to Him was in peace; and I was established by the Spirit of His government. Hallelujah.

\chapter{37}

\par \textit{An elementary Ode.}

\par 1 I stretched out my hands to my Lord: and to the Most I High I raised my voice:
\par 2 And I spake with the lips of my heart; and He heard me, when my voice reached I Him:
\par 3 His answer came to me, and gave me the fruits of my labours;
\par 4 And it gave me rest by the grace of the Lord. Hallelujah.



\chapter{38}

\par \textit{A beautiful description of the power of truth.}

\par 1 I went up to the light of truth as if into a chariot:
\par 2 And the Truth took me and led me: and carried me across pits and gulleys; and from the rocks and the waves it preserved me:
\par 3 And it became to me a haven of Salvation: and set me on the arms of immortal life:
\par 4 And it went with me and made me rest, and suffered me not to wander, because it was the Truth;
\par 5 And I ran no risk, because I walked with Him;
\par 6 And I did not make an error in anything because I obeyed the Truth.
\par 7 For Error flees away from it, and meets it not: but the Truth proceeds in the right path, and
\par 8 Whatever I did not know, it made clear to me, all the poisons of error, and the plagues of death which they think to be sweetness:
\par 9 And I saw the destroyer of destruction, when the bride who is corrupted is adorned: and the bridegroom who corrupts and is corrupted;
\par 10 And I asked the Truth ‘Who are these?’; and He said to me, 'This is the deceiver and the error:
\par 11 And they are alike in the beloved and in his bride: and they lead astray and corrupt the whole world:
\par 12 And they invite many to the banquet,
\par 13 And give them to drink of the wine of their intoxication, and remove their wisdom and knowledge, and so they make them without intelligence;
\par 14 And then they leave them; and then these go about like madmen corrupting: seeing that-they are without heart, nor do they seek for it!
\par 15 And I was made wise so as not to fall into the hands of the deceiver; and I congratulated myself because the Truth went with me,
\par 16 And I was established and lived and was redeemed,
\par 17 And my foundations were laid on the hand of the Lord: because He established me.
\par 18 For He set the root and watered it and fixed it and blessed it; and its fruits are for ever.
\par 19 It struck deep and sprung up and spread out, and was full and enlarged;
\par 20 And the Lord alone was glorified in His planting and in His husbandry: by His care and by the blessing of His lips,
\par 21 By the beautiful planting of His right hand: and by the discovery of His planting, and by the thought of His mind. Hallelujah.

\chapter{39}

\par \textit{One of the few allusions to events in the Gospels—that of our Lord walking on the Sea of Galilee.}

\par 1 Great rivers are the power of the Lord:
\par 2 And they carry headlong those who despise Him: and entangle their paths:
\par 3 And they sweep away their fords, and catch their bodies and destroy their lives.
\par 4 For they are more swift than lightning and more rapid, and those who cross them in faith are not moved;
\par 5 And those who walk on them without blemish shall not be afraid.
\par 6 For the sign in them is the Lord; and the sign is the way of those who cross in the name of the Lord;
\par 7 Put on, therefore, the name of the Most High, and know Him, and you shall cross without danger, for the rivers will be subject to you.
\par 8 The Lord has bridged them by His word; and He walked and crossed them on foot:
\par 9 And His footsteps stand firm on the water, and are not injured; they are as firm as a tree that is truly set up.
\par 10 And the waves were lifted up on this side and on that, but the footsteps of our Lord Messiah stand firm and are not obliterated and are not defaced.
\par 11 And a way has been appointed for those who cross after Him and for those who adhere to the course of faith in Him and worship His name. Hallelujah.

\chapter{40}

\par \textit{A song of praise without equal.}

\par 1 As the honey distills from the comb of the bees,
\par 2 And the milk flows from the woman that loves her children;
\par 3 So also is my hope on Thee, my God.
\par 4 As the fountain gushes out its water,
\par 5 So my heart gushes out the praise of the Lord and my lips utter praise to Him, and my tongue His psalms.
\par 6 And my face exults with His gladness, and my spirit exults in His love, and my soul shines in Him:
\par 7 And reverence confides in Him; and redemption in Him stands assured:
\par 8 And His inheritance is immortal life, and those who participate in it are incorrupt. Hallelujah.

\chapter{41}

\par \textit{We discover that the writer may be a Gentile (v. 8).}

\par 1 All the Lord's children will praise Him, and will collect the truth of His faith.
\par 2 And His children shall be known to Him. Therefore we will sing in His love:
\par 3 We live in the Lord by His grace: and life we receive in His Messiah:
\par 4 For a great day has shined upon us: and marvellous is He who has given us of His glory.
\par 5 Let us, therefore, all of us unite together in the name of the Lord, and let us honour Him in His goodness,
\par 6 And let our faces shine in His light: and let our hearts meditate in His love by night and by day.
\par 7 Let us exult with the joy of the Lord.
\par 8 All those will be astonished that see me. For from another race am I:
\par 9 For the Father of truth remembered me: He who possessed me from the beginning:
\par 10 For His bounty begat me, and the thought of His heart:
\par 11 And His Word is with us in all our way;
\par 12 The Saviour who makes alive and does not reject our souls-;
\par 13 The man who was humbled, and exalted by His own righteousness,
\par 14 The Son of the Most High appeared in the perfection of His Father;
\par 15 And light dawned from the Word that was beforetime in Him;
\par 16 The Messiah is truly one; and He was known before the foundation of the world,
\par 17 That He might save souls for ever by the truth of His name: a new song arises from those who love Him. Hallelujah.

\chapter{42}

\par \textit{The Odes of Solomon, the Son of David, are ended with the following exquisite verses.}

\par 1 I stretched out my hands and approached my Lord:
\par 2 For the stretching of my hands is His sign:
\par 3 My expansion is the outspread tree which was set up on the way of the Righteous One.
\par 4 And I became of no account to those who did not take hold of me; and I shall be with those who love me.
\par 5 All my persecutors are dead; and they sought after me who hoped in me, because I was alive:
\par 6 And I rose up and am with them; and I will speak by their mouths.
\par 7 For they have despised those who persecuted them;
\par 8 And I lifted up over them the yoke of my love;
\par 9 Like the arm of the bridegroom over the bride,
\par 10 So was my yoke over those that know me:
\par 11 And as the couch that is spread in the house of the bridegroom and bride,
\par 12 So is my love over those that believe in me.
\par 13 And I was not rejected though I was reckoned to be so.
\par 14 I did not perish, though they devised it against me.
\par 15 Sheol saw me and was made miserable:
\par 16 Death cast me up, and many along with me.
\par 17 I had gall and bitterness, and I went down with him to the utmost of his depth:
\par 18 And the feet and the head he let go, for they were not able to endure my face:
\par 19 And I made a congregation of living men amongst his dead men, and I spake with them by living lips:
\par 20 Because my word shall not be void:
\par 21 And those who had died ran towards me: and they cried and said, Son of God, have pity on us, and do with us according to thy kindness,
\par 22 And bring us out from the bonds of darkness: and open to us the door by which we shall come out to thee.
\par 23 For we see that our death has not touched thee.
\par 24 Let us also be redeemed with thee: for thou art our Redeemer.
\par 25 And I heard their voice; and my name I sealed upon their heads:
\par 26 For they are free men and they are mine. Hallelujah.


\end{document}