\begin{document}

\title{Greek Apocalypse of Ezra}

\chapter{1}

\par \textit{Introduction and prayer}

\par 1 It came to pass in the thirtieth year on the twenty-second of the month, I was in my house and I cried out, saying to the Most High, “Lord, grant (me) glory so

\par 2 [...]


\par 3 that I may see your mysteries.” When night fell the angel Michael, the archangel, came and said to me, “Prophet Ezra, lay aside bread for seventy weeks.” And

\par 4 I fasted just as he told me. And the archistrategos Raphael came and gave me a storax staff, and I fasted twice sixty weeks, and I saw the mysteries of God

\par 5 [...]

\par 6 and his angels. And I said to them, “I wish to plead with God concerning the Christian people. It were better that man were not born than that he entered the world.”

\par \textit{Ezra taken up to heaven: his prayer for mercy}

\par 7 Therefore, I was taken up into heaven and I saw in the first heaven a great

\par 8 command of angels and they led me to the judgments. And I heard a voice saying

\par 9 to me, “Have pity upon us, Ezra elect of God.” Then I began to say, “Woe to the sinners when they see the righteous man (elevated) above angels, and they are

\par 10 for fiery Gehenna.” And Ezra said, “Have pity upon the works of your hands,

\par 11 merciful and greatly pitying one. Condemn me rather than the souls of the sinners, for it is better to punish one soul and not to bring the whole world to

\par 12 destruction.” And God said, “I shall give rest to the righteous in Paradise and

\par 13 1 am merciful.” And Ezra said, “Lord, Why do you show favor to the righteous?

\par 14 For as a hired man completes his time of service and goes away, and again a slave serves his masters in order to receive his wage, thus the righteous man receives his reward in the heavens. But, have mercy upon the sinners for we know that

\par 15 [...]

\par 16 you are merciful.” And God said, “I have no way to be merciful to them.” And

\par 17 [...]

\par 18 Ezra said, “(Be merciful) because they cannot sustain your anger.” And God said, “(I am wrathful) because such (are the deserts) of such (men) as these.”

\par 19 And God said, “I wish to keep you as both Paul and John. You have given me uncorrupted the inviolate treasury, the treasure of virginity, the wall of men.”

\par 20 [...]

\par \textit{Ezra's second prayer}

\par 21 And Ezra said, “It were better if man were not born; it were well if he were

\par 22 not alive. The dumb beasts are a better thing than man, for they do not have

\par 23 punishment. You took us and delivered us to judgment. Woe to the sinners in the world to come, for their condemnation is endless and the flame unquenched.”

\chapter{2}

\par \textit{Ezra remonstrates with God: the sin of Adam}

\par 1 As I said this to him, Michael and Gabriel and all the apostles came and said,

\par 2 “Greetings!” lAnd Ezra said, “Faithful man of God! Arise and come hither with me, O Lord, to judgment.” And God said, “Behold I am giving you my

\par 3 [...]

\par 4 [...]

\par 5 covenant, both mine and yours, so that you will accept it.” And Ezra said, “We

\par 6 shall plead our case in your ear(s).” And God said, “Ask Abraham your father what kind of son presses suit against his father and come and plead the case with

\par 7 us.” And Ezra said, “As the Lord lives, I shall never cease pleading the case

\par 8 with you on account of the Christian people. Where are your former mercies, O

\par 9 Lord? Where your long-suffering?” And God said, “As I made night and day I made the righteous and the sinner and it were fitting to conduct yourself like the

\par 10 righteous man.” And the prophet said, “Who made Adam, the protoplast, the

\par 11 first one?” And God said, “My immaculate hands, and I placed him in Paradise

\par 12 to guard the region of the tree of life.” “Since he who established

\par 13 disobedience made this (man) sin.” And the prophet said, “Was he not guarded

\par 14 by an angel? And was life not preserved (by) the cherubim for the endless age?

\par 15 And how was he deceived who was guarded by angels (whom) you commanded

\par 16 to be present whatever happened? Attend also to that which I say! lf you had

\par 17 not given him Eve, the serpent would never have deceived her. If you save whom you wish you will also destroy whom you wish.”

\par \textit{Ezra remonstrates with God: the sins of men}

\par 18 And the prophet said, “O my Lord, let us continue to a second judgment.”

\par 19 And God said, “I cast fire upon Sodom and Gomorrah.” And the prophet said,

\par 20 [...]

\par 21 “Lord, you bring upon us what we deserve.” And God said, “Your sins exceed

\par 22 my kindness.” And the prophet said, “Remember Scripture, my father, who

\par 23 measured out Jerusalem and rebuilt her. Pity, Lord, the sinners, pity your own

\par 24 molding, have mercy upon your works.” Then God remembered his works and

\par 25 said to the prophet, “How can I have mercy upon them? They gave vinegar and gall to drink and [...] they repented.”

\par \textit{The day of judgment}

\par 26 And the prophet said, “Reveal yourcherubiml and let us go togethertojudgment,

\par 27 and show me what is the character of the day of judgment.” And God said, “You

\par 28 [...]

\par 29 have digressed, Ezra, for such is the day of judgment upon which there is no rain

\par 30 on the earth, for is a merciful judgment during that day”. And the prophet said, “I shall never cease to argue the case with you until I see the day of

\par 31 [...]

\par 32 consummation.” (And God said,) “Count the stars and the sand of the sea and if you will be able to count this, you will also be able to argue the case with me.”

\chapter{3}

\par 1 And the prophet said, “Lord, you know that I bear human flesh. And how can

\par 2 [...]

\par 3 I count the stars of heaven and the sand of the sea?” And God said, “O my elect prophet, no man will know that great day and the manifestation which prevails to

\par 4 judge the world. For your sake, O my prophet, I told you the day, but the hour

\par 5 I told you not.” And the prophet said, “Lord, tell me also the years.” And God said, “If I see that the justice of the world has become abundant, I will be long-suffering toward them. If not, I will stretch out my hand and I will grasp the inhabited world from its four corners and I will gather them all together to the valley of Jehosaphat and I will wipe out the human race and the world will be no

\par 6 [...]

\par 7 more.” And the prophet said, “And how will your right hand be glorified?”

\par 8 And God said, “I will be glorified by my angels.”

\par \textit{Why was man created?}

\par 9 And the prophet said, “Lord, if this was your calculation, Why did you form

\par 10 man? You said to Abraham our father, ‘I will surely multiply your seed as the stars of the heaven and as the sand along the shore of the sea.’ And where is your promise?”

\par \textit{Signs of the end}

\par 11 And God said, “First I shall cause by shakingd the fall of four-footed beasts and

\par 12 men. And when you see that brother delivers brother over to death and children

\par 13 will rise up against parents and a wife abandons her own husband, and when nation will rise up against nation in war, then you shall know that the end is near.

\par 14 And then brother will not have mercy upon brother, nor man upon his wife, nor

\par 15 children upon parents, nor friends upon friends, nor slave upon master. For the opponent of men himself will come up from Tartarus and will show many things

\par 16 to men. What shall I do to you, Ezra, and will you argue the case with me?”

\chapter{4}

\par \textit{Ezra descends to Tartarus}

\par 1 And the prophet said, “Lord, I shall never cease arguing the case with you.”

\par 2 And God said, “Count the flowers of the earth. If you can count them you also will be able to argue the case with me.”

\par 3 [...]

\par 4 And the prophet said, “Lord, I cannot count them—I bear human flesh—but

\par 5 neither will I stop arguing the case with you. I wish, Lord, to see the lower parts

\par 6 of Tartarus.” And God said, “Go down and see!” And he gave me Michael

\par 7 [...]

\par 8 and Gabriel and thirty-four other angels, and I descended eighty-five steps and they led me down five hundred steps.

\par \textit{The punishment of Herod}

\par 9 And I saw a fiery throne and an old man seated on it, and his punishment was

\par 10 merciless. And I said to the angels, “Who is this and what is his sin?” And they said to me, 

\par 11 “This is Herod, who was king for a time, and he commanded to kill

\par 12 the infants two years old or under.” And I said, “Woe upon his soul!” The disobedient and the abyss

\par 13 And again they led me down thirty steps. And I saw boiling fires there, and a

\par 14 multitude of sinners in them. And I heard their voices, but I did not perceive their

\par 15 forms. And they led me down deeper many steps which I was unable to count.

\par 16, And I saw old men there, and fiery axles were revolving upon their ears.d And

\par 17 [...]

\par 18 I said, “Who are these and what is their sin?” And they said to me, “These are

\par 19 the eavesdroppers.” And again they led me down five hundred Other steps. And

\par 20 [...]

\par 21 there I saw the unsleeping worm and fire consuming the sinners. And they led me down to the foundation of Apoleia (Destruction) and there I saw the twelvefold

\par 22 blow of the abyss. And they led me away to the south and there I saw a man

\par 23 hanging from his eyelids and the angels were beating him. And I asked, “Who

\par 24 is this and what is his sin?” And Michael the archistrategos said to me, “This man is incestuous; having carried out a small lust, this man was commanded to be hanged.”

\par \textit{The Antichrist}

\par 25 And they led me away to the north and I saw a man there restrained with iron

\par 26 bars. And I asked, “Who is this?” And he said to me, “This is the one who says, ‘I am the son of God and he who made stones bread and water wine.’”

\par 27 [...]

\par 28 And the prophet said, “Make known to me what sort of appearance he has and

\par 29 I will inform the race of men lest they believe in him.” And he said to me, “The appearance of his face is as of a wild man. His right eye is like a star rising at

\par 30 dawn and the other is unmoving. His mouth is one cubit, his teeth are a span

\par 31 long, his fingers like scythes, the soles of his feet two span, and on his forehead

\par 32 an inscription ‘Antichrist.’ He was exalted up to heaven, he will descend as far

\par 33 as Hades. One time he will be a child, another an old man.” And the prophet

\par 34 [...]

\par 35 said, “Lord, how do you permit the race of men to stray?” And God said, “Hear, my prophet! He becomes a Child and an old man and let no one believe him that

\par 36 he is my beloved son. And after these things a trumpet, and the graves will be

\par 37 opened and the dead will rise up uncorrupted. Then the opponent, having heard

\par 38 the terrible threat, will hide himself in the outer darkness. Then the heaven and

\par 39 the earth and the sea will perish. Then I shall bum the heaven for eighty cubits

\par 40 and the earth for eight hundred cubits.” And the prophet said, “And (in) what

\par 41 did the heaven sin?” And God said, “Since [...] is the evil.” And the prophet

\par 42 [...]

\par 43 said, “Lord, (in) what did the earth sin?” And God said, “Since the opponent having heard my terrible threat will hide (in it), and because of that I shall melt the earth and with it the rebel of the race of men.”

\chapter{5}

\par \textit{Further punishments}

\par 1 And the prophet said, “Pity, O Lord, the race of Christians.” And I saw a

\par 2 [...]

\par 3 woman suspended and four wild beasts were sucking upon her breasts. And the angels said to me, “She begrudged giving her milk but also cast infants into the 

\par 4 rivers.” And I saw terrible darkness and night without stars or moon. There is there neither young nor old, neither brother with brother nor mother with Child nor

\par 5 [...]

\par 6 wife with husband. And I wept and said, “O Lord, Lord, have mercy upon the sinners.”

\par \textit{Ezra taken to heaven}

\par 7 And as I said these things a cloudb came and seized me and took me up again

\par 8 to the heavens. And I saw many judgments and I wept bitterly and I said, “It

\par 9 [...]

\par 10 were better if man did not come forth from his mother's belly.” Those who were in punishment called out, saying, “Since you came here, holy one of God, we

\par 11 have obtained a slight respite.” And the prophet said, “Blessed are they who bewail their own sins.”

\par \textit{Birth and its purpose}

\par 12 And God said, “Hear Ezra, beloved one! Just as a farmer casts down the seed

\par 13 of corn into the earth, so a man casts down his seed into a woman's place. In the first (month) it is a whole, in the second it is swollen, in the third it grows hair, in the fourth it grows nails, in the fifth it becomes milky, in the sixth it is ready and quickened, in the seventh it is prepared, [in the eighth...], in the ninth the bars of the gateways of the woman are opened and it is born healthy on the earth.”

\par 14 And the prophet said, “It were better for man not to have been born. Alas, O

\par 15 [...]

\par 16 human race, at that time when you come to judgment!” And I said to the Lord,

\par 17 “Lord, Why did you create man and give him over to judgment?” And God said in his exalted pronouncement, “I will not pardon those who transgress my

\par 18 covenant.” And the prophet said, “Lord, where is your goodness?” And God said, I prepared everything because of man and man does not keep my commandments.

\par 19 [...]

\par \textit{Punishments and rewards}

\par 20 And the prophet said, “Lord, reveal to me the punishments and Paradise.”

\par 21 And the angels led me away to the eastf and I saw the tree of life. And I saw there 

\par 22 Enoch and Elijah and Moses and Peter and Paul and Luke and Matthew and all

\par 23 the righteous and the patriarchs.g And I saw there the rpunishmentl of the air and the blowing of the winds and the storehouses of the ice and the eternal

\par 24 punishments. And I saw there a man hanging by his skull. And they said to me,

\par 25 [...]

\par 26 “This one transferred boundaries.” And there I saw great judgments and said to

\par 27 the Lord, “O Lord, Lord, which of men, having been born, did not sin?” And they led me farther down in Tartarus and I saw all the sinners lamenting and

\par 28 weeping and evil mourning. And I too wept, seeing the race of rnen punished thus.

\chapter{6}

\par 1 Then God said to me, “Ezra, do you know the names of the angels who are

\par 2 over the consummation: Michael, Gabriel, Uriel, Raphael, Gabuthelon, Aker, Arphugitonos, Beburos, Zebuleon?”

\par \textit{Ezra struggles for his soul}

\par 3 Then a voice came to me, “Come here, die, Ezra, my beloved! Give back that

\par 4 which has been entrusted (to you).” And the prophet said, “And whence can

\par 5 you bring forth my soul?” And the angels said, “We can cast it forth through

\par 6 your mouth.” And the prophet said, “I spoke mouth to mouth with God and it

\par 7 will not go forth from there.” And the angels said. “We will bring it forth through

\par 8 your nostrils.” And the prophet said, “My nostrils smelled the glory of God.”"

\par 9 And the angels said, “We can bring it forth through your eyes.” And the prophet

\par 10 [...]

\par 11 said, “My eyes have seen the back of God.” And the angels said, “We can bring

\par 12 it forth through your head.” And the angels said, “We can bring it forth through your feet.” And the prophet said, “I walked with Moses on the mountain,

\par 13 and it will not come forth from there.” And the angels said, “We can cast it

\par 14 forth through the tips of your (toe)nails.” And the prophet said, “My feet walked in in the sanctuary.” And the angels departed unsuccessful, saying, “Lord, we

\par 15 [...]

\par 16 cannot receive his soul.” Then he said to his only begotten son, “Go down, my beloved son, with a numerous host of angels, taking the soul of my beloved Ezra.”

\par 17 For the Lord, having taken a numerous army of many angels, said to the prophet, “Give me that deposit which I entrusted to you. The crown is readied for you.”

\par 18 And the prophet said, “Lord, if you take my soul from me, who will you have

\par 19 left to plead on behalf of the race of men?” And God said, “You who are mortal

\par 20 and earthly, do not plead the case with me.” And the prophet said, “I shall never

\par 21 cease pleading.” And God said, “Give, in the meanwhile, that which is entrusted

\par 22 (to you). The crown is readied for you. Come here, die, so that you may attain

\par 23 it.” Then the prophet began to speak with tears, “O Lord, what profits it that I

\par 24 plead the case with you, and I am going to fall to the earth? Woe, woe! for I will

\par 25 be consumed by worms. Bewail me, all holy and pious ones, I plead greatly and

\par 26 am delivered over to death! Bewail me, all holy and just ones, because I have entered the bowl of Hades.”

\chapter{7}

\par \textit{Soul and body}

\par 1 And God said to him, “Hear, Ezra, my beloved one. I, being immortal, received

\par 2 a cross, I tasted vinegar and gall, I was set down in a grave. And I raised up my elect ones and I summoned up Adam from Hades so that the race of men

\par 3 [...]. Therefore, fear not death. For that which is from mes that is the soul, departs for heaven. That which is from the earth, that is the body, departs for the

\par 4 earth from which it was taken.” And the prophet said, “Woe, woe! What shall I do? How shall I act? I know not.”

\par \textit{Concluding prayerb}

\par 5 And then the blessed Ezra began to say, “O eternal God, Creator of the whole creation, who measured out the heaven with a span and contained the earth in his

\par 6 hand, who drives the cherubim, who took the prophet Elijah to the heavens in

\par 7 a fiery chariot, who gives nurture to all flesh, whom all things fear and tremble

\par 8 from the face of your power, hear me who pleads greatly and give to all who 

\par 9 copy this book and preserve it and recall my name and preserve my memory fully,

\par 10 give them blessing from heaven. And bless all of his things, just as the ends of

\par 11 Joseph.d And remember not his previous sins on the day of his judgment. Those

\par 12 [...]

\par 13 who do not believe this book will be burned like Sodom and Gomorrah.” And a voice carne to him saying, “Ezra. my beloved, I shall grant to each one the things which you asked.”

\par \textit{Death and burial of Ezra}

\par 14 And at once he gave over his precious soul with much honor on the eighteenth

\par 15 of the month of October. And they buried him with incense and psalms. His precious and holy body provides unceasingly strengthening of souls and bodies for those who approach him willingly.

\par 16 Glory, might, honor, and worship to him for whom it is fitting, for the Father and the Son and the Holy Spirit, now and always and for ever and ever. Amen.

\end{document}