\begin{document}

\title{Revelation of Ezra}

\chapter{1}

\par textit{The revelation which was made to Ezra and the children of Israel concerning the nature of the year througha the beginning of January.}

\par 1 If the first day of January comes on the Lord's Day, it makes a warm winter, a wet spring, a windy autumn, good crops, abundance of cattle, sufficient honey, good vintages, plenty of beans, successful gardens. (But) young men will die, there will be battles and great robberies, (and) something new will be heard about kings and rulers.

\par 2 The Day of the Moon (Monday) makes both winter (and) summer moderate. There will be great floods and sickness, infantry warfare, changes of rulers, many wives will sit in lamentation, there will be much ice, kings will die, (there will be) a good vintage, bees will die.

\par 3 The Day of Mars (Tuesday) makes a severe and gloomy winter, spring snow, rainy summer, (and) a dry autumn. Grain will be high-priced. (There will be) a destruction of pigs, (and) a sudden plague among cattle. Sailing (will be) dangerous, (and) honey sufficient; flax (will be) high-priced, fires (will be) numerous, beans, garden vegetables, (and) oil will be abundant. Women will die and kings (also will die). The vintage (will be) troubled.

\par 4 The Day of Mercury (Wednesday): productiveness of crops, a good vintage, lack of fruits, success in business, a destruction of men, a warm winter. Autumn will be moderate. (There will be) dangers from the sword, plenty of oil, looseness of the bowels and entrails. Women will die, there will be famine in diverse places (and) a good summer. Something new will be heard (and) there will be no honey.

\par 5 The Day of Jupiter (Thursday): (There will be) worthlessness of grain; meat (will be) high-priced (and there will be an) abundance of fruits. There will be less honey; winter (will be) moderate, spring windy, autumn good, summer good. (There will be) a destruction of pigs (and) much rain; rivers will flood.f There will be suficient oil, the crop will spoil, beans will be mixed,g and (there will be) peace.

\par 6 The Day of Venus (Friday) makes a moderate winter, a bad summer, dry autumn, worthless grain, a good vintage, inflammation of the eyes. Infants will die, there will be an earthquake, (there will be) peril for kings; oil (will be) abundant, sheep and bees will perish.

\par 7 The Day of Saturn (Saturday) makes a windy winter, a bad spring, a summer (which) changes as it is harassed by storms, a dry autumn, scarcity of grain (and) high-priced flax. Fevers will spread,h people will be harassed by various ailments, old men will die.


\end{document}